---
title: "Giant Frogs in Bend"
author: "Corey Dow-Hygelund"
date: "1/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Loading

The dataset is contained in an Excel (.xlsx) spreadsheet, located in the subdirectory `data`. The raw statistics are on the sheet _All Female raw statistics_. Lets load it in and take a quick look.
```{r data_load, message=FALSE}
library(readxl)

file_path <- 'data/Female size statistics.2.xlsx'
data_df <- read_excel(file_path, range=(cell_cols('B:G')), 
                      sheet="All Females raw statistics")

head(data_df)
```

## Summary statistics

Before plotting, let's take a look at summary numbers. This helps ensure that the plots are being correctly rendered.

There are a total of `r nrow(data_df)` observations, for `r ncol(data_df)` sites. 


```{r summary}
summary(data_df)
```

Sunriver (SR) has significantly more observations than the other sites, with the second highest Dilman Meadows (DM) containing only half the measurements. This will influence plots based upon distributions.  

## Plotting

The data is currently in the so-called _wide_ format, where each value of the site location variable is an independent column. Let's tranform it to the _long_ format, where each observation is contained in a single column, with another single column denoting site location.

In addition, there are many `NA` (NULL) values, due to the difference in number of observations across sites. These are easily removed with `na.omit`. 

```{r reshape}
library(tidyr)
data_long = na.omit(gather(data_df, site_location, size, 
                           CR:LSA, factor_key=TRUE))
head(data_long)
```

There are now `r nrow(data_long)` unique observations for _all_ sites.

### Violin

Let's take an initial stab at producing a Violin plot. 

```{r violin_1}
library(ggplot2)
p <- ggplot(data_long, aes(x=site_location, y=size)) + 
#   geom_violin(scale='count')
  geom_violin()
p + geom_boxplot(width=0.1)

```

