---
title: "Giant Frogs in Bend"
author: 
  - "Corey Dow-Hygelund,  Mozilla"
  - "Tlell Wolf, Bend Science Station"
  - "Jay Bowerman, Sunriver Nature Center"
date: "1/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Loading

The dataset is contained in an Excel (.xlsx) spreadsheet, located in the subdirectory `data`. The raw statistics are on the sheet _All Female raw statistics_. Lets load it in and take a quick look.
```{r data_load, message=FALSE}
library(readxl)

file_path <- 'data/Female size statistics.2.xlsx'
data_df <- read_excel(file_path, range=(cell_cols('B:G')), 
                      sheet="All Females raw statistics")

head(data_df)
```

## Summary statistics

Before plotting, let's take a look at summary numbers. This helps ensure that the plots are being correctly rendered.

There are a total of `r nrow(data_df)` observations, for `r ncol(data_df)` sites. 


```{r summary}
summary(data_df)
```

Sunriver (SR) has significantly more observations than the other sites, with the second highest Dilman Meadows (DM) containing only half the measurements. This will influence plots based upon distributions.  

## Plotting

The data is currently in the so-called _wide_ format, where each value of the site location variable is an independent column. Let's tranform it to the _long_ format, where each observation is contained in a single column, with another single column denoting site location.

In addition, there are many `NA` (NULL) values, due to the difference in number of observations across sites. These are easily removed with `na.omit`. 

```{r reshape}
library(tidyr)
data_long = na.omit(gather(data_df, site_location, size, 
                           CR:LSA, factor_key=FALSE))
head(data_long)
```

There are now `r nrow(data_long)` unique observations for _all_ sites.

<<<<<<< HEAD
### Violin/Boxplot
=======
### Violing/Boxplot
>>>>>>> a60988acdba9f41701c930b8a928523f5b2f4e44

```{r violin_1}
library(ggplot2)
p <- ggplot(data_long, aes(x=site_location, y=size)) +
  labs(x = 'Location', y = 'Length (mm)') +
  theme_classic() + 
  theme(axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"))
#   geom_violin(scale='count')
p +  
  geom_violin(draw_quantiles = c(0.5)) 

```

Standard boxplot 
```{r boxplot}
p + geom_boxplot()
```


```{r boxplot_violin}
p +  
  geom_violin() +
  geom_boxplot(width=0.1)
  
```

Producing the violin plot, where the violin area is scaled by the number of counts

```{r violin_2}
p +  
  geom_violin(scale='count')

```

### Ridge

Rather than representing the di
```{r ridge_plot}
library(ggridges)
p <- ggplot(data_long, aes(x=size, y=site_location)) + 
  labs(x = 'Length (mm)', y = 'Location') +
  theme_classic() + 
  theme(axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"))
p + geom_density_ridges()
```

Reducing the overlap between the histograms to zero.

```{r}
p + geom_density_ridges(scale = 1)
```

Adding in quantiles

```{r}
p + geom_density_ridges(scale = 1, quantile_lines=TRUE, quantiles=2)
```

#### Adding Color
Color representing probability of distribution

```{r}
ggplot(data_long, aes(x=size, y=site_location, fill=0.5 - abs(0.5-..ecdf..))) + 
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  labs(x = 'Length (mm)', y = 'Location') +
  theme_classic() + 
  theme(axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"))
```

Color representing length 
```{r}
ggplot(data_long, aes(x=size, y=site_location, fill=..x..)) + 
  geom_density_ridges_gradient() +
  scale_fill_viridis_c(name = "Temp. [F]", option = "C")
  labs(x = 'Length (mm)', y = 'Location') +
  theme_classic() + 
  theme(axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"))
  
```

